### 2018-06-22 dissertation/1-intro/chapter1.tex
Updated preface and chapter 1

In 2013, an estimated 1.25 million people lost their lives due to traffic accidents globally \citep{whotraffic}.
While this number has greatly reduced over the past decades \citep{johnson2010traffic} thanks to better safety features (seat belts, air bags, ABS, etc.) and stronger laws (drunk driving, motorcycle helmets, speed limits, etc.), it may still rise as a major cause of death in the years to come \citep{whofactsheet}, so further actions are necessary.
Furthermore, as the car population increases, congestions consume ever more time of the daily commuter, peaking at over 100 hours per year for the residents of Los Angeles, CA \citep{inrixtraffic}.

@@

@@

These issues — the additional safety and efficiency as well as the low-latency communications — can be tackled through the use of a vehicular ad-hoc network (VANET), in which vehicles share data amongst themselves without relying on external devices, an Internet connection or server availability.
Neighboring vehicles can share their position and velocity data at high frequencies, allowing, for example, for autonomous vehicles to plan a platooning approach to traffic \citep{amoozadeh2015platoon}.
In the case of a collision or other event, nearby nodes can broadcast alerts, which other nodes pick up and forward \citep{li2007routing}.
That way, an alert can travel long distances in little time, allowing approaching vehicles to safely slow down or pick alternative routes.

@@

%\section{Document organization}
The remainder of this document is organized as follows. \Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in technological and social networks.
Then, \autoref{chap:vanets} goes into details regarding VANETs and the importance of trust solutions in the field, presenting previous studies made on the subject.
\Cref{chap:truman} describes the goals of TruMan, its underlying hypothesis and the work done to achieve those goals, as well as the previously existing algorithms which form parts of TruMan.
Next, \autoref{chap:simulations} explains the tools used to validate TruMan's functionality and presents the results of the experiments that were performed.
Finally \autoref{chap:conclusion} presents the final thoughts on the project and concludes this document.

%Finally, \autoref{chap:proposal} shows the fundamentals of this proposed work, including the similarities found between VANETs and social networks, a realistic movement model that may be used for simulations, the previously existing study on malicious node identification for complex networks, and the important aspects that must be adapted to the dynamic vehicular environment.

>>>>

In 2013, an estimated 1.25 million people lost their lives due to traffic accidents globally \citep{whotraffic}.
While this number has greatly reduced over the past decades \citep{johnson2010traffic} \citep{iihsfatality} thanks to better safety features (seat belts, air bags, ABS, etc.) and stronger laws (drunk driving, motorcycle helmets, speed limits, etc.), it may still rise as a major cause of death in the years to come \citep{whofactsheet}, so further actions are necessary.
Furthermore, as the car population increases, congestions consume even more time of the daily commuter, peaking at over 100 hours per year for the residents of Los Angeles, CA \citep{inrixtraffic}.
Moscow, New York, Bogotá, São Paulo, London and Paris are also among the ten most congested cities in the world \citep{inrixtraffic}.

@@

Today, certain vehicle manufacturers already include the on-board technology required to enable vehicular communications in the real world \citep{toyota2015}.
However, this technology only becomes truly useful when there are other vehicles or infrastructure with whom to communicate, so, at the moment, the benefits are notable but limited \citep{cadillac2017}.

@@

These issues — the additional safety and efficiency as well as the low-latency communications — can be tackled through the use of a vehicular ad-hoc network (VANET), in which vehicles share data amongst themselves without relying on external devices, an Internet connection or server availability.
Neighboring vehicles can share their position and velocity data at high frequencies, allowing, for example, for autonomous vehicles to plan a platooning approach to traffic \citep{amoozadeh2015platoon}.
In the case of a collision or other event, nearby nodes can broadcast alerts, which other nodes pick up and forward \citep{li2007routing}, as illustrated by \Cref{fig:collision}.
That way, an alert can travel long distances in little time, allowing approaching vehicles to safely slow down or pick alternative routes.

@@

%\section{Document organization}
The remainder of this document is organized as follows. \Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in technological and social networks.
Then, \Cref{chap:vanets} goes into details regarding VANETs and the importance of trust solutions in the field, presenting previous studies made on the subject.
\Cref{chap:truman} describes the goals of TruMan, its underlying hypothesis and the work done to achieve those goals, as well as the previously existing algorithms which form parts of TruMan.
Next, \Cref{chap:simulations} explains the tools used to validate TruMan's functionality and presents the results of the experiments that were performed.
Finally \Cref{chap:conclusion} presents the final thoughts on the project and concludes this document.

%Finally, \Cref{chap:proposal} shows the fundamentals of this proposed work, including the similarities found between VANETs and social networks, a realistic movement model that may be used for simulations, the previously existing study on malicious node identification for complex networks, and the important aspects that must be adapted to the dynamic vehicular environment.

### 2018-06-22 dissertation/2-complexnetworks/chapter2.tex
Updated preface and chapter 1

\chapter{Trust in Complex Networks}
\label{chap:complexnetworks}

@@

\begin{enumerate}
	\item \textbf{Technological Networks} are grids purposefully engineered to provide services to consumers and/or citizens.
	 	The primary examples of these networks are the Internet, the telephone network, power grids, transportation and delivery networks.
	 	A commonly studied type of technological network are Mobile Ad-hoc Networks (MANETs).
		Although not of widespread use, MANETs can provide a way to create a network without pre-existing infrastructure, as long as each device is equipped with the proper hardware and software.
		Trust issues in technological networks and MANETs are detailed in \autoref{section:trusttechnological}.
		VANETs, which are special types of MANETs, are introduced in \autoref{chap:vanets}, along with several details regarding trust in those types of networks.
	\item \textbf{Social Networks} are formed of relationships between people, or groups of people.
		These relationships can be familiar, friendships, acquaintance, etc.
		For the purposes of this work, the most relevant type of relationship is that of trust.
		The details surrounding trust relationships in social networks are shown in \autoref{section:trustsocial}.
	\item \textbf{Information Networks} are the ones in which nodes are pieces of data or information and the edges are the connections between those pieces.
		Often, information networks are directly associated with technological or social networks.
		For instance, while the World Wide Web is an information network (in which the nodes are webpages and the edges are the links that users click on to navigate), it relies on the Internet, as it contains the physical infrastructure that makes the web possible.
		Online social networks can also be classified as information networks, since their nodes are actually information about people rather than the people themselves.
		Trust in information networks can be observed in some instances, like peer-to-peer networks, although its usages are not relevant for this work.
	\item \textbf{Biological Networks} are the networks found in nature.
		Their nodes can be chemicals, cells, animals, groups of lifeforms, and more.
		An example is the brain, which contains a neural network formed by neurons; connections in the network represent signals that are sent from one neuron to another.
		Another instance of biological networks are food chains, categorized as ecological webs.
		Species of animals are the nodes, while the predation of other species form the edges.
		In biological networks, it is difficult to clearly define trust, since nodes may not have any sort of awareness or intelligence (such as cells or proteins).
		Regardless, the study of trust in biological networks is not relevant for this proposal.

@@

In the following two sections of this chapter, trust is further explained in the contexts of social and technological networks.
Both types of network can be fitted into the model above, but contain distinct features that demand closer examination.
Furthermore, features of both are relevant when analyzing network structure and trust graphs for vehicular networks, which are expanded upon in \autoref{chap:vanets}.

@@

In general, social networks' topology and trust graphs are mostly static.
Although friendships are formed and ended frequently (i.e. the topology is dynamic), those connections do not disrupt the general shape of the network, because members of the network will usually have other friends whose relationships remain stable.
Even if a certain person's trust integrity is compromised due to a specific incident, that person's friends are not necessarily deemed untrustworthy, preserving part of the trust graph.
While positive trust is often tied to the social topology, it is not always the case: one example is two work colleagues who may have a professional relationship, but wouldn't trust each other on other matters; another is the trust people place in authority figures without necessarily having met. 
\autoref{fig:social_network} shows an example of a small social network containing a family and an office.

In \autoref{section:socialvanets}, the argument is made that VANETs can be considered social networks in several occasions and how this can be used to develop trust in vehicular networks.

@@

% In ad-hoc networks, such as vehicular networks, which are detailed in \autoref{chap:vanets}, trust can be handled in very different ways.
While the centralized trust solutions above serve their purpose on the security and privacy of Internet users, they would be too slow to be viable in a dynamic ad-hoc network, which cannot rely on a back-end infrastructure to distribute those lists.
The most common instance of mobile ad-hoc networks, or MANETs, is using mobile devices, such as smartphones, being carried by humans.
Although these networks are dynamic, their mobility is relatively low in relation to the wireless range of the devices — if two people are walking in opposite directions, their phones may communicate for several seconds before they leave each other's range.
Trust solutions for MANETs can use this property to their advantage, since it allows one node to test another and check several of the messages sent between them.

@@

\autoref{fig:manet} is an example highlighting the difference between the topology graph and the trust graph.
The topology graph changes as one node moves across the network, with edges being added and removed as nodes move into and out of each other's ranges.
The trust graph, however, is constantly being built, maintaining past relationships even if the nodes are no longer in communication range.

>>>>

\chapter{Complex Networks and Trust}
\label{chap:complexnetworks}

@@

\begin{enumerate}
	\item \textbf{Technological Networks} are grids purposefully engineered to provide services to consumers and/or citizens.
	 	The primary examples of these networks are the Internet, the telephone network, power grids, transportation and delivery networks.
	 	A commonly studied type of technological network are Mobile Ad-hoc Networks (MANETs).
		Although not of widespread use, MANETs can provide a way to create a network without pre-existing infrastructure, as long as each device is equipped with the proper hardware and software.
		Trust issues in technological networks and MANETs are detailed in \Cref{section:trusttechnological}.
		VANETs, which are special types of MANETs, are introduced in \Cref{chap:vanets}, along with several details regarding trust in those types of networks.
	\item \textbf{Social Networks} are formed of relationships between people, or groups of people.
		These relationships can be familiar, friendships, acquaintance, etc.
		For the purposes of this work, the most relevant type of relationship is that of trust.
		The details surrounding trust relationships in social networks are shown in \Cref{section:trustsocial}.
	\item \textbf{Information Networks} are the ones in which nodes are pieces of data or information and the edges are the connections between those pieces.
		Often, information networks are directly associated with technological or social networks.
		For instance, while the World Wide Web is an information network (in which the nodes are webpages and the edges are the links that users click on to navigate), it relies on the Internet, as it contains the physical infrastructure that makes the web possible.
		Online social networks can also be classified as information networks, since their nodes are actually information about people rather than the people themselves.
		Trust in information networks can be observed in some instances, like peer-to-peer networks, although its usages are not relevant for this work.
	\item \textbf{Biological Networks} are the networks found in nature.
		Their nodes can be chemicals, cells, animals, groups of lifeforms, and more.
		An example is the brain, which contains a neural network formed by neurons; connections in the network represent signals that are sent from one neuron to another.
		Another instance of biological networks are food chains, categorized as ecological webs.
		Species of animals are the nodes, while the predation of other species form the edges.
		In biological networks, it is difficult to clearly define trust, since nodes may not have any sort of awareness or intelligence (such as cells or proteins).
		Regardless, the study of trust in biological networks is not relevant for this proposal.

@@

In the following two sections of this chapter, trust is further explained in the contexts of social and technological networks.
Both types of network can be fitted into the model above, but contain distinct features that demand closer examination.
Furthermore, features of both are relevant when analyzing network structure and trust graphs for vehicular networks, which are expanded upon in \Cref{chap:vanets}.

@@

In general, social networks' topology and trust graphs are mostly static.
Although friendships are formed and ended frequently (i.e. the topology is dynamic), those connections do not disrupt the general shape of the network, because members of the network will usually have other friends whose relationships remain stable.
Even if a certain person's trust integrity is compromised due to a specific incident, that person's friends are not necessarily deemed untrustworthy, preserving part of the trust graph.
While positive trust is often tied to the social topology, it is not always the case: one example is two work colleagues who may have a professional relationship, but wouldn't trust each other on other matters; another is the trust people place in authority figures without necessarily having met. 
\Cref{fig:social_network} shows an example of a small social network containing a family and an office.

In \Cref{section:socialvanets}, the argument is made that VANETs can be considered social networks in several occasions and how this can be used to develop trust in vehicular networks.

@@

% In ad-hoc networks, such as vehicular networks, which are detailed in \Cref{chap:vanets}, trust can be handled in very different ways.
While the centralized trust solutions above serve their purpose on the security and privacy of Internet users, they would be too slow to be viable in a dynamic ad-hoc network, which cannot rely on a back-end infrastructure to distribute those lists.
The most common instance of mobile ad-hoc networks, or MANETs, is using mobile devices, such as smartphones, being carried by humans.
Although these networks are dynamic, their mobility is relatively low in relation to the wireless range of the devices — if two people are walking in opposite directions, their phones may communicate for several seconds before they leave each other's range.
Trust solutions for MANETs can use this property to their advantage, since it allows one node to test another and check several of the messages sent between them.

@@

\Cref{fig:manet} is an example highlighting the difference between the topology graph and the trust graph.
The topology graph changes as one node moves across the network, with edges being added and removed as nodes move into and out of each other's ranges.
The trust graph, however, is constantly being built, maintaining past relationships even if the nodes are no longer in communication range.

### 2018-06-22 dissertation/3-vanets/chapter3.tex
Updated chapters 1 and 2
Updated preface and chapter 1

\chapter{Vehicular Ad-hoc Networks}
\label{chap:vanets}

@@

%The advantages of an ad-hoc approach, as opposed to using cellular infrastructure or other server-based solutions, are explained in \autoref{chap:introduction}.
%For messages to travel to nodes out of the sender's wireless range, they must use other nodes as \textit{hops}.
%This approach requires appropriate routing protocols that consider the dynamic topology of VANETs.
%While traditional MANETs often use protocols which rely on the network topology, the ones proposed for VANETs generally use geographic position instead \citep{saini2015close}; these protocols use GPS coordinates in order to send messages to the available neighbor physically closest to the destination node.

@@

Each of these possible attacks requires a unique approach, though there are some broader ways to help the security and safety of VANET users.
Trust, as is described in \autoref{section:trustvanet}, can be an important feature in vehicular networks, especially when attempting to filter out malicious or incorrect messages.
It does not, however, avoid all possible attacks, such as a signal jammer or stalking.
Rather, different mechanisms must be explored in order to avoid most problems.

@@

The analysis of related work is based on \citep{zhang2011survey}, which proposes eight desired properties for a trust management model for VANETs.
In this section, these properties are briefly described.
% with an assessment of whether or not TruMan satisfies their conditions.
%\autoref{table:properties} shows how well TruMan and previous work satisfy the properties.

@@

%Although the model proposed here will not take role-based trust into consideration and does not emphasize event-related messages, the author's approach to experience-based trust resembles what will be described in \autoref{chap:proposal}.

@@

%conclude
\autoref{table:properties} shows how well each studied trust model satisfies the eight properties described in \autoref{section:properties}.
The numbers used in the table are the same ones from \autoref{section:properties}.
This analysis shows that, although several previously proposed trust models provide reasonable solutions to the trust management problem, all of them have issues in terms of correctness, completeness or efficiency, indicating that the problem is not yet truly solved.

>>>>

\chapter{Trust in Vehicular Ad-hoc Networks}
\label{chap:vanets}

@@

%The advantages of an ad-hoc approach, as opposed to using cellular infrastructure or other server-based solutions, are explained in \Cref{chap:introduction}.
%For messages to travel to nodes out of the sender's wireless range, they must use other nodes as \textit{hops}.
%This approach requires appropriate routing protocols that consider the dynamic topology of VANETs.
%While traditional MANETs often use protocols which rely on the network topology, the ones proposed for VANETs generally use geographic position instead \citep{saini2015close}; these protocols use GPS coordinates in order to send messages to the available neighbor physically closest to the destination node.

@@

Each of these possible attacks requires a unique approach, though there are some broader ways to help the security and safety of VANET users.
Trust, as is described in \Cref{section:trustvanet}, can be an important feature in vehicular networks, especially when attempting to filter out malicious or incorrect messages.
It does not, however, avoid all possible attacks, such as a signal jammer or stalking.
Rather, different mechanisms must be explored in order to avoid most problems.

@@

The analysis of related work is based on \citep{zhang2011survey}, which proposes eight desired properties for a trust management model for VANETs.
In this section, these properties are briefly described.
% with an assessment of whether or not TruMan satisfies their conditions.
%\Cref{table:properties} shows how well TruMan and previous work satisfy the properties.

@@

%Although the model proposed here will not take role-based trust into consideration and does not emphasize event-related messages, the author's approach to experience-based trust resembles what will be described in \Cref{chap:proposal}.

@@

%conclude
\Cref{table:properties} shows how well each studied trust model satisfies the eight properties described in \Cref{section:properties}.
The numbers used in the table are the same ones from \Cref{section:properties}.
This analysis shows that, although several previously proposed trust models provide reasonable solutions to the trust management problem, all of them have issues in terms of correctness, completeness or efficiency, indicating that the problem is not yet truly solved.

### 2018-06-22 dissertation/4-project/chapter4.tex
updates to keynote and minor updates to dissertation
Updated preface and chapter 1

%This work shows that it is possible to adapt a trust management scheme originally designed for static networks to a dynamic environment, such as a VANET.
This work introduces an efficient solution to trust management in dynamic networks such as VANETs.
In order to make this possible, it is necessary to identify features in VANETs that show that nodes can share a long-term relationship, as is the case for social networks.
%The arguments for such features are presented in \autoref{section:socialvanets}, while the movement model that enables the simulation of the features is explained in \autoref{section:workingday}.
Through these long-term relationships, it then becomes feasible for nodes to store trust data and share it with other nodes.
By combining a node's own opinions about familiar nodes and trust information received from its neighbors, it is possible to create a model of the surrounding network.
This model includes a trust graph, showing the trust relationships between pair of nodes, which can then be used in conjunction with other algorithms in order to classify nodes as correct or malicious.

@@

Furthermore, the desired properties of a VANET trust model described in \citep{zhang2011survey} (explained in detail in \autoref{section:properties}) were considered, so TruMan attempts to fulfill or enable as many of those properties as possible.

@@

The number of components is, at most, $|V|$: in a worst-case scenario, each node is placed into its own component.
The complexity of the algorithm is $O(|V|+|E|)$ for a graph $T = (V,E)$.
Algorithm \autoref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \citep{tarjan1972depth}.

@@

\autoref{fig:tarjan} illustrates the execution of Tarjan's algorithm.
The algorithm starts from node 0, with $index[0] = 0$ and $lowlink[0] = 0$.
With a depth-first search, the algorithm traces the path $0 \rightarrow 1 \rightarrow 2 \rightarrow 3 
\rightarrow 4 \rightarrow 2$.
Since node $2$ has already been visited and nodes 4 and 3 have no further outgoing edges, $lowlink[4]$ and $lowlink[3]$ receive the value 2 and the function calls return back to the first visit of node $2$.
At this point, nodes 2, 3 and 4 all have 2 as the smallest reachable index and, therefore, they form a strongly connected component.

@@

The complexity of the algorithm is $O(|E'|)$ for a graph $C = (V',E')$.
As a comparison, the DSATUR algorithm for graph coloring has complexity $O(|V|^2)$ \citep{brelaz1979new}.
Algorithm \autoref{algorithm:coloring} shows the general structure of the graph coloring algorithm  \citep{mittal2011graph}.

@@

\autoref{fig:coloring} illustrates the execution of the graph coloring algorithm.
It is notable how the algorithm takes few iterations to fully color the graph.
However, it is also possible to observe that the result does not use the minimum amount of colors.
By coloring node 4 as cyan and node 3 as magenta, the sample graph could have been colored with only three colors instead of four.
As described above, this is not a problem for the usage of the algorithm in TruMan.

@@

First, $T$ is separated into strongly connected components using Tarjan's algorithm \citep{tarjan1972depth}, which is described in detail in \autoref{section:tarjan}.
In each of these components, all nodes are connected by edges of value $1$.
In other words, within a single component, all nodes trust one another; nodes connected by edges of value $0$ are separated into different components.
Each of these components becomes a node of a component graph $C = (V', E')$.

@@

In the experiments performed by the authors of MaNI, the coloring heuristic shows the most promising results, identifying a high ratio of the malicious nodes in the network.
The coloring heuristic uses a graph coloring algorithm, such as DSATUR \citep{brelaz1979new} or the algorithm detailed in \autoref{section:coloring}.
Other heuristics were experimented with, but were either less effective in detecting malicious nodes, provided too many false positives, or were not efficient enough.

@@

\autoref{fig:mani} illustrates the execution of the MaNI algorithm.
With the starting graph $T$, whose edges represent the trust relationships between nodes, Tarjan's strongly connected component algorithm is executed.
\autoref{fig:mani1} shows nodes colored according to their placement in a strongly connected component.
The strongly connected components form a graph $C$ according to edges present in $T$.
In \autoref{fig:mani2}, component $0$ is the one containing node $5$; component $1$ contains nodes $6$, $7$ and $8$; component $2$ contains nodes $0$, $1$, $2$ and $3$; and component $3$ contains node $4$.
The graph coloring with minimum colors algorithm is executed on $C$, producing the coloration shown in \autoref{fig:mani3}.
Finally, each node in $T$ is colored according to which color its component received in $C$.
The color with the most nodes is deemed benign, while the others are considered malicious.

@@

Every time a neighboring node $v$ is tested as benign, the value of $u \rightarrow v$ increases.
Additionally, node $u$ performs an union between its trust graph and $v$'s trust graph, forming a new graph $T^u_i = T^u_{i-1} \bigcup T^v_{i-1}$, which is used for the remaining steps.
Algorithm  \autoref{algorithm:interaction} shows the basic interaction between two nodes, while algorithm \autoref{algorithm:union} details the steps of the graph union.

\begin{algorithm}
\caption{Interaction between two nodes}\label{algorithm:interaction}
\begin{algorithmic}[1]

\Function{Interaction}{node $u$, node $v$}

@@

\If{$v$ is benign}
	\State $T^u(u\rightarrow v).trustvalue$ \textbf{increases}
	\State \textbf{Union}($T^u$, $T^v$)
\Else
	\State $T^u(u\rightarrow v).trustvalue$ \textbf{decreases}
\EndIf

@@

@@

	\If{$a\notin T^u$}
		\State \textbf{add} $a$ \textbf{to} $T^u$
	\EndIf

	\If{$b\notin T^u$}
		\State \textbf{add} $b$ \textbf{to} $T^u$
	\EndIf

	\If{$a\rightarrow b \notin T^u$}
		\State \textbf{add} $a\rightarrow b$ \textbf{to} $T^u$
	\EndIf
\EndFor

@@

The coloring heuristic is used to classify nodes, which uses the algorithm described in \autoref{section:coloring} \citep{mittal2011graph}, although other heuristics may be considered.
After running the graph coloring algorithm with graph $C^u_i$ as input, the color whose nodes in $C^u_i$ represent the most nodes in $T^u_i$ is classified as correct, and all others are classified as malicious.
Once this information from $C^u_i$ is brought back to graph $T^u_i$, it is trivial to label the nodes in $T^u_i$ as either benign or malicious based on the classifications of their components.

@@

Then, when node $u$ interacts with another node $w$ in the future, $u$'s opinion of $v$ comes with the timestamp attached, so $w$ has the information to know how much time has passed since the interaction between $u$ and $v$ happened.
Once $w$ performs the graph merge procedure with $T_u$ and $T_w$, but $T_w$ already has an edge $u\rightarrow v$ stored, it checks the two timestamps and only updates the edge if the incoming information is more recent than the node already had stored.
Since the timestamp is always set to the destination node of the edge, it can be used for comparison regardless of which nodes handled the information. 
Algorithm \autoref{algorithm:union_timestamp} shows how the union function was changed to accommodate timestamps and information aging.

@@

	\If{$a\notin T^u$}
		\State \textbf{add} $a$ \textbf{to} $T^u$
	\EndIf

	\If{$b\notin T^u$}
		\State \textbf{add} $b$ \textbf{to} $T^u$
	\EndIf

	\If{$a\rightarrow b \notin T^u$}
		\State \textbf{add} $a\rightarrow b$ \textbf{to} $T^u$
	\Else
		\State $s_0 \gets T^u(a\rightarrow b).timestamp$
		\State $s_1 \gets T^v(a\rightarrow b).timestamp$
		\If{$s_1$ \textbf{is more recent than} $s_0$}
			\State $T^u(a\rightarrow b).trustvalue \gets T^v(a\rightarrow b).trustvalue$
			\State $T^u(a\rightarrow b).timestamp \gets T^v(a\rightarrow b).timestamp$
		\EndIf	
	\EndIf
\EndFor

@@

Information aging is necessary because, when a node converts from benign to malicious, old information might still indicate that it is benign, as shown in \autoref{fig:rogue}.
Immediately after node $0$ becomes malicious, it is still part of a benign strongly connected component, because there are paths from every other node to $0$, and from $0$ to every other node.
Some time later, node $2$ interacts with $0$ again and identifies it as malicious; other nodes that interact with $2$ will receive that updated information.
However, the outdated edge $3\rightarrow 0$ keeps $0$ in the component, since other nodes trust $3$ and $3$ still hasn't updated its own opinion of $0$.
Once that edge is old enough, it is discarded and node $0$ is correctly identified as malicious by all other nodes.

>>>>

%This work shows that it is possible to adapt a trust management scheme originally designed for static networks to a dynamic environment, such as a VANET.
This work introduces an efficient solution to trust management in dynamic networks such as VANETs.
In order to make this possible, it is necessary to identify features in VANETs that show that nodes can share a long-term relationship, as is the case for social networks.
%The arguments for such features are presented in \Cref{section:socialvanets}, while the movement model that enables the simulation of the features is explained in \Cref{section:workingday}.
Through these long-term relationships, it then becomes feasible for nodes to store trust data and share it with other nodes.
By combining a node's own opinions about familiar nodes and trust information received from its neighbors, it is possible to create a model of the surrounding network.
This model includes a trust graph, showing the trust relationships between pair of nodes, which can then be used in conjunction with other algorithms in order to classify nodes as correct or malicious.

@@

Furthermore, the desired properties of a VANET trust model described in \citep{zhang2011survey} (explained in detail in \Cref{section:properties}) were considered, so TruMan attempts to fulfill or enable as many of those properties as possible.

@@

The number of components is, at most, $|V|$: in a worst-case scenario, each node is placed into its own component.
The complexity of the algorithm is $O(|V|+|E|)$ for a graph $T = (V,E)$.
Algorithm \Cref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \citep{tarjan1972depth}.

@@

\Cref{fig:tarjan} illustrates the execution of Tarjan's algorithm.
The algorithm starts from node 0, with $index[0] = 0$ and $lowlink[0] = 0$.
With a depth-first search, the algorithm traces the path $0 \rightarrow 1 \rightarrow 2 \rightarrow 3 
\rightarrow 4 \rightarrow 2$.
Since node $2$ has already been visited and nodes 4 and 3 have no further outgoing edges, $lowlink[4]$ and $lowlink[3]$ receive the value 2 and the function calls return back to the first visit of node $2$.
At this point, nodes 2, 3 and 4 all have 2 as the smallest reachable index and, therefore, they form a strongly connected component.

@@

The complexity of the algorithm is $O(|E'|)$ for a graph $C = (V',E')$.
As a comparison, the DSATUR algorithm for graph coloring has complexity $O(|V|^2)$ \citep{brelaz1979new}.
Algorithm \Cref{algorithm:coloring} shows the general structure of the graph coloring algorithm  \citep{mittal2011graph}.

@@

\Cref{fig:coloring} illustrates the execution of the graph coloring algorithm.
It is notable how the algorithm takes few iterations to fully color the graph.
However, it is also possible to observe that the result does not use the minimum amount of colors.
By coloring node 4 as cyan and node 3 as magenta, the sample graph could have been colored with only three colors instead of four.
As described above, this is not a problem for the usage of the algorithm in TruMan.

@@

First, $T$ is separated into strongly connected components using Tarjan's algorithm \citep{tarjan1972depth}, which is described in detail in \Cref{section:tarjan}.
In each of these components, all nodes are connected by edges of value $1$.
In other words, within a single component, all nodes trust one another; nodes connected by edges of value $0$ are separated into different components.
Each of these components becomes a node of a component graph $C = (V', E')$.

@@

In the experiments performed by the authors of MaNI, the coloring heuristic shows the most promising results, identifying a high ratio of the malicious nodes in the network.
The coloring heuristic uses a graph coloring algorithm, such as DSATUR \citep{brelaz1979new} or the algorithm detailed in \Cref{section:coloring}.
Other heuristics were experimented with, but were either less effective in detecting malicious nodes, provided too many false positives, or were not efficient enough.

@@

\Cref{fig:mani} illustrates the execution of the MaNI algorithm.
With the starting graph $T$, whose edges represent the trust relationships between nodes, Tarjan's strongly connected component algorithm is executed.
\Cref{fig:mani1} shows nodes colored according to their placement in a strongly connected component.
The strongly connected components form a graph $C$ according to edges present in $T$.
In \Cref{fig:mani2}, component $0$ is the one containing node $5$; component $1$ contains nodes $6$, $7$ and $8$; component $2$ contains nodes $0$, $1$, $2$ and $3$; and component $3$ contains node $4$.
The graph coloring with minimum colors algorithm is executed on $C$, producing the coloration shown in \Cref{fig:mani3}.
Finally, each node in $T$ is colored according to which color its component received in $C$.
The color with the most nodes is deemed benign, while the others are considered malicious.

@@

Every time a neighboring node $v$ is tested as benign, the value of $u \rightarrow v$ increases.
Additionally, node $u$ performs an union between its trust graph and $v$'s trust graph, forming a new graph $T^u_i = T^u_{i-1} \bigcup T^v_{i-1}$, which is used for the remaining steps.
Algorithm  \Cref{algorithm:interaction} shows the basic interaction between two nodes, while algorithm \Cref{algorithm:union} details the steps of the graph union.

\begin{algorithm}
\caption{Interaction between two nodes $u$ and $v$}\label{algorithm:interaction}
\begin{algorithmic}[1]

@@

\If{$v$ is benign}
	\State $T^u(u\rightarrow v).trustvalue$ \textbf{increases}
	\State $T^u \gets $ \textbf{Union}($T^u$, $T^v$)
\Else
	\State $T^u(u\rightarrow v).trustvalue$ \textbf{decreases}
\EndIf

\end{algorithmic}
\end{algorithm}

@@

\State $T \gets T^u$

@@

	\If{$a\notin T$}
		\State \textbf{add} $a$ \textbf{to} $T$
	\EndIf

	\If{$b\notin T$}
		\State \textbf{add} $b$ \textbf{to} $T$
	\EndIf

	\If{$a\rightarrow b \notin T$}
		\State \textbf{add} $a\rightarrow b$ \textbf{to} $T$
	\EndIf

\EndFor

\State \textbf{return} $T$

@@

The coloring heuristic is used to classify nodes, which uses the algorithm described in \Cref{section:coloring} \citep{mittal2011graph}, although other heuristics may be considered.
After running the graph coloring algorithm with graph $C^u_i$ as input, the color whose nodes in $C^u_i$ represent the most nodes in $T^u_i$ is classified as correct, and all others are classified as malicious.
Once this information from $C^u_i$ is brought back to graph $T^u_i$, it is trivial to label the nodes in $T^u_i$ as either benign or malicious based on the classifications of their components.

@@

Then, when node $u$ interacts with another node $w$ in the future, $u$'s opinion of $v$ comes with the timestamp attached, so $w$ has the information to know how much time has passed since the interaction between $u$ and $v$ happened.
Once $w$ performs the graph merge procedure with $T_u$ and $T_w$, but $T_w$ already has an edge $u\rightarrow v$ stored, it checks the two timestamps and only updates the edge if the incoming information is more recent than the node already had stored.
Since the timestamp is always set to the destination node of the edge, it can be used for comparison regardless of which nodes handled the information. 
Algorithm \Cref{algorithm:union_timestamp} shows how the union function was changed to accommodate timestamps and information aging.

@@

	\If{$a\notin T$}
		\State \textbf{add} $a$ \textbf{to} $T$
	\EndIf

	\If{$b\notin T$}
		\State \textbf{add} $b$ \textbf{to} $T$
	\EndIf

	\If{$a\rightarrow b \notin T$}
		\State \textbf{add} $a\rightarrow b$ \textbf{to} $T$
	\Else
		\State $s_u \gets T(a\rightarrow b).timestamp$
		\State $s_v \gets T^v(a\rightarrow b).timestamp$
		\If{$s_u$ \textbf{is more recent than} $s_v$}
			\State $T(a\rightarrow b).trustvalue \gets T^v(a\rightarrow b).trustvalue$
			\State $T(a\rightarrow b).timestamp \gets T^v(a\rightarrow b).timestamp$
		\EndIf	
	\EndIf
\EndFor

\State \textbf{return} $T$

@@

Information aging is necessary because, when a node converts from benign to malicious, old information might still indicate that it is benign, as shown in \Cref{fig:rogue}.
Immediately after node $0$ becomes malicious, it is still part of a benign strongly connected component, because there are paths from every other node to $0$, and from $0$ to every other node.
Some time later, node $2$ interacts with $0$ again and identifies it as malicious; other nodes that interact with $2$ will receive that updated information.
However, the outdated edge $3\rightarrow 0$ keeps $0$ in the component, since other nodes trust $3$ and $3$ still hasn't updated its own opinion of $0$.
Once that edge is old enough, it is discarded and node $0$ is correctly identified as malicious by all other nodes.
